---
title: "Problem Set 4"
author: "Sarah Stonedahl"
date: "3/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

# Submission

1. Fork this repository to your own account
1. Make changes to the `README.Rmd` file (including the author field above).
1. Knit the file to HTML format
1. Publish HTML on Rpubs (you will need to sign up for the free account) and *add the link below*.
1. Submit a PR with your answers.

*Link to Rpubs document:* http://rpubs.com/stonedas/268933

# Overview

Take the [datacamp course on joining data](https://www.datacamp.com/courses/joining-data-in-r-with-dplyr) to learn the join verbs.

You will analyze some data in the [`nihexporter`](https://github.com/jayhesselberth/nihexporter) package to answer some questions. The questions will cover methods in `dplyr` including two-table verbs.

First you need to install the `nihexporter` package (and `devtools`, if you don't have it already).

```r
# install.packages('devtools')
devtools::install_github("jayhesselberth/nihexporter")
```

Read the [Overview vignette](http://jayhesselberth.github.io/nihexporter/articles/nihexporter.html) to learn the structure of the pacakge. Take a look at two tables from the package:

```{r overview}
library(nihexporter)
library(tidyverse)
library(broom)
```

# Problems

## Problem 1

The `projects` and `project_io` tables have the `r intersect(names(projects), names(project_io))` column in common. Use this column to perform a `left_join` and identify the most expensive project for each institute. Project institutes are in the `projects` table, whereas total project costs are available in the `project_io` table. Note you will have to use the `arrange` and `slice` verbs from `dplyr`.

```{r problem_1, eval=FALSE}

left_join(project_io, projects, by = "project.num") %>% arrange(desc(total.cost)) %>% group_by(institute) %>% slice(1) -> most_expensive_projects

most_expensive_projects %>% arrange(desc(total.cost)) -> ordered_most_expensive_projects

ordered_most_expensive_projects
```

## Problem 2

Create a plot that describes the change in output (number of publications) for the most expensive grant from the above list. Note that you cannot use `n.pubs` for this, as that is an aggregate count over time. You need a tbl of following structure, which you can then use to calculate the number of publications each year (hint: use `n()` in a `summarize` statement:

```{r p2_tbl}
left_join(project_io, projects, by = "project.num") %>% arrange(desc(total.cost)) %>% group_by(institute) %>% slice(1) -> most_expensive_projects

most_expensive_projects %>% arrange(desc(total.cost)) -> ordered_most_expensive_projects

ordered_most_expensive_projects %>% head(1) %>% left_join(publinks) %>% left_join(publications) %>% select(project.num, pub.year, fy.cost, total.cost, pmid) %>% na.omit() %>% select(project.num, institute, total.cost, pub.year, pmid) %>% group_by(project.num, institute, total.cost, pub.year) %>% summarize(total.pubs= n()) %>% na.omit() -> p2_tbl

p2_tbl
```

```{r problem 2}
ggplot(p2_tbl, aes(pub.year, total.pubs)) + geom_col(fill = "darkblue") + xlab('Publication Year') + ylab('Total Publications') -> p2_plot 

p2_plot
```

## Problem 3

Use the `projects` and `clinical_studies` table to calculate how many clinical studies have been started by projects from each institute.

```{r problem_3}
select(projects, institute, project.num) %>% left_join(clinical_studies) %>% na.omit() %>% select(institute, project.num, trial.id) %>% group_by(institute, project.num) %>% summarise(total.studies = n()) %>% arrange(desc(total.studies)) -> p3_tbl

p3_tbl
```

## Problem 4

Calculate correlations between `total.cost` and `n.pubs` for each institute. You will need to `join` the `projects` and `project_io` tables to link `institute` with `total.cost` and `n.pubs`. You will then need to use `group_by` and `do` from `dplyr` and `tidy` from `broom` to calculate correlations.

```{r problem_4, eval=FALSE}
projects %>% left_join(project_io) %>% select(institute, total.cost, n.pubs) %>% na.omit() %>% group_by(institute) %>% summarize(total_cost = sum(total.cost), num_pubs = sum(n.pubs)) %>% do(tidy(cor(.$total_cost, .$num_pubs))) -> correlation_cost_pubs

correlation_cost_pubs
```

## Problem 5 (extra credit)

Correlate [`rcr`](https://icite.od.nih.gov/) values from the `publications` table to one of:

- total cost (overall for NIH, or by institute)
- budget mechanism (e.g., RFA, PAR, etc.)
- type of grant (e.g. R01, P01, R21, etc.). 

```{r extra_credit}
```

## Problem 6 (extra extra credit)

Use [`gganimate`](https://github.com/dgrtwo/gganimate) to illustrate the change in cost of R01s for a few institutes over time.

```{r extra_extra_credit}
```
